Learning Auditory Similarities Between Instrumental Playing Techniques

### 2018-01-29 paper/lagrangeSpontaneousSimilarity2017.tex
ml answers to vl comments

\author{
Christian El-Hajj,
Mathias Rossignol,
Gr\'egoire Lafay,
Mathieu Lagrange}

@@

\vl{The paragraph above is extremely vague. I don't think that "opportunistic" is the word we're looking for here. @writing}

@@

\vl{It should be made very clear that the IPT taxonomy is
the set product of the instrument taxonomy
and the playing technique taxonomy. @writing}

@@

\vl{We should probably say a word on the Hornbostel-Sachs taxonomy,
which is based on player-instrument interaction, but is not outdated
given the variety of contemporary music techniques. @biblio}

@@

\vl{The paragraph above needs to be rewritten. It introduces mathematical
notations that are not used later in the text. The adjective "plagued"
is derogatory and could be replaced by "hindered". @writing}

In order to reduce the number of items, we decided to perform a selection
experiment where the subject is asked to give his opinion on which
\ipts is relevant to study. An \ipt is said to be relevant to study
if it is likely to be associated to another \ipt of another instrument.
Interest is rated on a 7 ticks scale. The subject can listen to all the different
nuance and pitch samples of each \ipt. The experiment is over when all the
\ipts have been rated.

@@

Two music composition professors of the
"conservatoire national sup\'erieur de musique et de danse de Paris",
a renowned French school of music, performed the test.
All the couples with a grade higher than three are selected,
bringing 78 \ipts to study.

\vl{Too much passive form in this subsection. This last paragraph ("a renowned French school of music") is un-academic. @writing}

@@

@@

\begin{figure}
\center
\includegraphics[width = \textwidth]{figures/xp2example.png}
\caption{Spatial and group organization provided by subject 1.
\vl{
Why have you chosen subject 1, who made 14 clusters?
This is an unusually large number, according to the
histogram in Figure 2.
It would be more representative of the population to illustrate this figure
with the data of a subject with the median number of clusters.
It would be good to show the three stages of the human-computer interface:
random initialization,
geometric clustering on screen,
and color clustering.
Besides, the figure is not accessible to readers with
color vision deficiencies at the moment.
The color of the clusters should be reinforced with
geometrical information, \eg{} Voronoi cells.
@data @figure}}
\label{fig:xp2display}
\end{figure}

@@

\begin{figure}
\center
\includegraphics[width = \textwidth]{figures/nbc.png}
\caption{Histogram of the number of groups.
\vl{The width of this figure should be a half page instead of a full page. Pair with next figure. @figure}}
\label{fig:xp2nbGroup}
\end{figure}

\begin{figure}
\center
\includegraphics[width = \textwidth]{figures/sbc.png}
\caption{Histogram of the size of groups.
\vl{The width of this figure should be a half page instead of a full page. Pair with previous figure. @figure}}
\label{fig:xp2sizeGroup}
\end{figure}

@@

\begin{figure}
\center
\includegraphics[width = \textwidth]{figures/clusterAnalysis.png}
\caption{Cluster analysis of the similarity judgments. $j/I$ and $j/Pt$ compare the partition obtained using the similarity judgments clustered using a growing number of clusters versus respectively the instrument partition and the playing technique partition. Normalizing null partitions based NMI $nullI$ and $nullPt$ are used to obtain normalized measures: $nj/I$ and $nj/Pt$.
\vl{This line plot is very hard to understand.
The legend box occludes the data on the top-right corner.
It would be preferable to remove this legend box and label lines
by placing text upon them,
so that the plot would be legible to readers with color vision deficiencies.
Besides, the choice of line labels is cryptic, as it relies on abbreviations.
The distinction between I and Pt
can be illustrated by two complementary colors, \eg{} blue versus orange.
The distinction between j, nj, and null
can be illustrated by three distinct markers
(empty circle, filled triangle, asterisk),
associated with
three distinct line styles (solid, dashed, dotted).
The NMI abbreviation should be expanded.
@figure}}
\label{fig:clusters}
\end{figure}

@@

\begin{figure}
\center
\includegraphics[width = \textwidth]{figures/groupModes.png}
\caption{Clustering \ipts using average perceptual similarity into 5 clusters (top) and playing technique labels (top). The color chosen to display each clusters on top corresponds to the playing technique that is dominant in this cluster, from left to right: 'pizzicato', 'slap', 'key', 'ordinario', 'ordinario', 'ponticello', 'art'.
\vl{A bubble plot would be preferable to a color map. Abbreviated words should appear in their expanded form instead. @writing}}
\label{fig:gm}
\end{figure}

\begin{figure}
\center
\includegraphics[width = \textwidth]{figures/groupInstruments.png}
\caption{Clustering \ipts using average perceptual similarity into 8 clusters (a) and instrument reference labels (b). The color chosen to display the groups (on top) corresponds  to the instrument that is dominant in this group, from left to right: 'ASax', 'BBTb', 'Vc', 'ASax', 'TpC', 'Gtr', 'ASax', 'Vc'.
\vl{A bubble plot would be preferable to a color map. Abbreviated words should appear in their expanded form instead. @writing}}
\label{fig:gi}
\end{figure}

@@

\vl{Naming the two stages "primary" and "secondary" is not very informative.
How about "feature extraction" and "similarity learning"? @writing}

@@

\vl{The argument behind not training a deep convolutional network is a strawman argument.
There is a also learning stage inherent to our approach: LDA or LMNN. @writing}

\vl{I do believe that time-frequency scattering (or one of its variants, see McDermott or Patil or Sch\"{a}dler)
is, in a very precise way, mandatory to model the central auditory system. It's OK to think otherwise but that
should be motivated. @writing}

\vl{I would suggest trying to avoid the invention of the "TST" acronym for "temporal scattering transform" if possible.
This acronym does not already exist in the relevant literature. @writing}

@@

\vl{I am a bit wary about not giving proper names to the two experiments, but instead calling them "experiment 1" and "experiment 2".
How about "expert judgments of uniqueness" versus "crowdsourced judgments of similarity"? @writing}

@@

\vl{The paragraph above is the most important in the whole paper, and it deserves more attention. How do these three solutions differ? Which one was chosen in the end? Did it make a significant difference in NMI? Is the consensus clustering shown somewhere in the paper? Is it related to the dendrogram in Appendix B? I have so many questions. @writing}

\vl{I'm afraid there is also a methodological problem here.
We need to have a hold-out set of IPTs to evaluate the performance of the clustering algorithm after having trained the metric learning algorithm.
@methodology @computation}

@@

\vl{What does "safely" mean here? We are at the epicenter of the paper here: every word matters. @writing}

@@

As explained in the first section, gathering perceptual similarity data is plagued with a size problem. Thus, the perceptual data considers 78 \ipts each represented by one sound recording, most often of C4 pitch and nuance mezzo forte. In order to provide enough data for the numerical analysis described in this section, we complement the data gathered during the second experiment as follows: for each \ipt, instead of one sound exemplar we consider the whole range of nuance and pitch that is available. This leads to a dataset of 9346 sound recordings. It is thus assumed that the perceptual similarity is invariant with respect to pitch and nuance which is not necessarily true, but is taken in this paper as a reasonable assumption.
\vl{It's too bad that this "reasonable" assumption is not challenged against the data. In music psychology, the "bandwidth for timbre invariance" is of the order of one octave (Handel and Erickson, Music Perception, 2001) or less (Steele and Williams, Music Perception, 2006).
One way to do it without collecting new similarity judgments would be to break down the precision by pitch of query.
Presumably we have an inverted U shape with a peak near C4.
@research @computation @figure @biblio}

@@

\vl{I'm extremely puzzled by the paragraph above. I thought that the point of the paper was precisely to liberate oneself from the taxonomy of musical instruments? @writing}

@@

\vl{The mel scale has a varying quality factor, so I don't see how this can be right.}
\vl{I understand what "logarithmically compressed" and "normalized using the median" because I did work on this before but,
for someone who is less experienced, this might not be precise enough.}

@@

\vl{The paragraph above is very peremptory and at the same time quite vague.
I could quote many tasks in which there is no need to "target in some way" the scattering transform to obtain better results than MFCC.
What scientific conclusion are we drawing here? @writing}

@@

\begin{table}
  \caption{Precision achieved by the different settings. Numbers in bold face indicate best performance. \vl{What does these variances correspond to? Variance across raters, across IPTs, across instances of each IPT? Furthermore, why are the results with TST, LMNN, and "reference" commented? Do they correspond to same rater between training and evaluation? @methodology @writing}}
  \label{tab:res1}
  \begin{center}
\begin{tabular}{lllc}
features & projection & reference & precision (\%) \\
  \hline
  MFCCs & - & - &   85.07 $\pm$ 6.19 \\
  MFCCs & LDA &  consensus  &   81.50 $\pm$ 7.65 \\
MFCCs & LMNN & multiple & 86.31 $\pm$ 5.91 \\
MFCCs & LMNN & consensus &   86.22 $\pm$ 5.92 \\
% MFCCs & LMNN & reference &   86.18 $\pm$ 6.05 \\
TST & - & - &   87.01 $\pm$ 5.81 \\
TST & LDA & consensus &   80.95 $\pm$ 10.37 \\
TST & LMNN  & multiple  &   93.31 $\pm$ 3.92 \\
TST & LMNN & consensus &   \textbf{94.80 $\pm$ 3.26} \\
% TST & LMNN & reference &   98.09 $\pm$ 1.28 \\
\end{tabular}
\end{center}
\end{table}

@@

\begin{figure}
\center
\includegraphics[width = \textwidth]{figures/frame.png}
\caption{Spatial and group organization provided by subject 1.
\vl{This caption is obviously wrong, as it was copied verbatim from Figure 1. The x axis should be made logarithmic, which uneven tick marks.
The legend is difficult to understand (acronyms, subtle difference between "multiple" and "consensus").
Again, the legend box could be removed after placing the labels next to the lines themselves. @figure}}
\label{fig:frame}
\end{figure}

@@

For the sake of reproducibility, the code of the proposed method and the experimental protocol is available\footnote{https://github.com/mathieulagrange/paperSpontaneousSimilarity \vl{I would love a more memorable name. @writing}}. For convenience, the features (scattering and MFCCs) are also provided, but the software can be used to process audio data.

@@

Thanks to the digitization capabilities available, gathering large amount of audio recordings of \ipt is achievable and several databases are now available on the market. Browsing of such databases can be achieved by keyword-based search. It requires that 1) the database is fully annotated with a precise ontology of instruments and playing techniques, 2) that the user has a precise understanding of this ontology, and 3) that the user a precise idea of the kind of timbre he is looking to find the \ipt he is searching for.

\vl{Curiously, this paper does not contain a review of such databases available on the market. @biblio}

\vl{Let's use the gender-neutral singular "they" instead of "he".@writing}

@@

Some instruments can be complemented with sordina (S), wha (W). The addition of this kind of device modifying the shape of the instrument, a saxophone alto and saxophone alto with sordina are considered as different instruments.
\vl{This is crucial and should not be deferred to the appendix. @writing}

@@

\vl{This section is not precise enough in terms of intensity ranges and pitch ranges.
The abbreviations can be removed. @data @writing}

\vl{A bubble plot of number of notes per instrument (x axis) and playing technique
(y axis) would be extremely useful. @data @figure}

@@

\vl{This dendrogram is very hard to read. Abbreviations must be replaced by the full names of every instrument and playing techniques.
Another way to visualize it would be a two-dimensional embedding, such as t-SNE or Isomap (I'm biased towards Isomap). @figure}

>>>>

\author{
Christian El-Hajj,
Vincent Lostanlen,
Mathias Rossignol,
Gr\'egoire Lafay,
Mathieu Lagrange}

@@

\vl{The paragraph above is extremely vague. I don't think that "opportunistic" is the word we're looking for here. @writing}
\ml{ok}

@@

\vl{It should be made very clear that the IPT taxonomy is
the set product of the instrument taxonomy
and the playing technique taxonomy. @writing}
\ml{ok}

@@

\vl{We should probably say a word on the Hornbostel-Sachs taxonomy,
which is based on player-instrument interaction, but is not outdated
given the variety of contemporary music techniques. @biblio}
\ml{great, I was missing such references}

@@

\vl{The paragraph above needs to be rewritten. It introduces mathematical
notations that are not used later in the text. The adjective "plagued"
is derogatory and could be replaced by "hindered". @writing}
\ml{ok}

In order to reduce the number of items, we perform a selection
experiment where the subject is asked to give his opinion on which
\ipts is relevant to study. An \ipt is said to be relevant to study
if it is likely to be associated to another \ipt of another instrument.
Interest is rated on a 7 ticks scale. The subject can listen to all the different
nuance and pitch samples of each \ipt. The experiment is over when all the
\ipts are rated.

@@

Two music composition professors of the
"conservatoire national sup\'erieur de musique et de danse de Paris", performed the test.
All the couples with a grade higher than three are selected,
bringing 78 \ipts to study.

@@

\ml{ok, je regarde si on a toujours les données}

@@

\begin{figure}
\center
\includegraphics[width = \textwidth]{figures/xp2example.png}
\caption{Spatial and group organization provided by subject 1.
\vl{
Why have you chosen subject 1, who made 14 clusters?
This is an unusually large number, according to the
histogram in Figure 2.
It would be more representative of the population to illustrate this figure
with the data of a subject with the median number of clusters.
It would be good to show the three stages of the human-computer interface:
random initialization,
geometric clustering on screen,
and color clustering.
Besides, the figure is not accessible to readers with
color vision deficiencies at the moment.
The color of the clusters should be reinforced with
geometrical information, \eg{} Voronoi cells.
@data @figure}}
\ml{ok, tu peux utiliser le site de demo et faire une capture d'écran de l'utilisateur qui te va. Pour les autres display, je ne suis pas sur que cela soit utile.}
\label{fig:xp2display}
\end{figure}

@@

\begin{figure}
\center
\includegraphics[width = \textwidth]{figures/nbc.png}
\caption{Histogram of the number of groups.
\vl{The width of this figure should be a half page instead of a full page. Pair with next figure. @figure}}
\ml{ok}
\label{fig:xp2nbGroup}
\end{figure}

\begin{figure}
\center
\includegraphics[width = \textwidth]{figures/sbc.png}
\caption{Histogram of the size of groups.
\vl{The width of this figure should be a half page instead of a full page. Pair with previous figure. @figure}}
\ml{ok}
\label{fig:xp2sizeGroup}
\end{figure}

@@

\begin{figure}
\center
\includegraphics[width = \textwidth]{figures/clusterAnalysis.png}
\caption{Cluster analysis of the similarity judgments. $j/I$ and $j/Pt$ compare the partition obtained using the similarity judgments clustered using a growing number of clusters versus respectively the instrument partition and the playing technique partition. Normalizing null partitions based NMI $nullI$ and $nullPt$ are used to obtain normalized measures: $nj/I$ and $nj/Pt$.
\vl{This line plot is very hard to understand.
The legend box occludes the data on the top-right corner.
It would be preferable to remove this legend box and label lines
by placing text upon them,
so that the plot would be legible to readers with color vision deficiencies.
Besides, the choice of line labels is cryptic, as it relies on abbreviations.
The distinction between I and Pt
can be illustrated by two complementary colors, \eg{} blue versus orange.
The distinction between j, nj, and null
can be illustrated by three distinct markers
(empty circle, filled triangle, asterisk),
associated with
three distinct line styles (solid, dashed, dotted).
The NMI abbreviation should be expanded.
@figure}}
\ml{ok, le .fig est dispo}
\label{fig:clusters}
\end{figure}

@@

\begin{figure}
\center
\includegraphics[width = \textwidth]{figures/groupModes.png}
\caption{Clustering \ipts using average perceptual similarity into 5 clusters (top) and playing technique labels (top). The color chosen to display each clusters on top corresponds to the playing technique that is dominant in this cluster, from left to right: 'pizzicato', 'slap', 'key', 'ordinario', 'ordinario', 'ponticello', 'art'.
\vl{A bubble plot would be preferable to a color map. Abbreviated words should appear in their expanded form instead. @writing}}
\ml{Je demande a voir, mais je ne suis pas satisfait de cette visu de toute facon. Le code est dispo dans le rep code. Lancer ticelSim en premier.}
\label{fig:gm}
\end{figure}

\begin{figure}
\center
\includegraphics[width = \textwidth]{figures/groupInstruments.png}
\caption{Clustering \ipts using average perceptual similarity into 8 clusters (a) and instrument reference labels (b). The color chosen to display the groups (on top) corresponds  to the instrument that is dominant in this group, from left to right: 'ASax', 'BBTb', 'Vc', 'ASax', 'TpC', 'Gtr', 'ASax', 'Vc'.
\vl{A bubble plot would be preferable to a color map. Abbreviated words should appear in their expanded form instead. @writing}}
\ml{ok}
\label{fig:gi}
\end{figure}

@@

\vl{Naming the two stages "primary" and "secondary" is not very informative.
How about "feature extraction" and "similarity learning"? @writing}
\ml{ok}

@@

\vl{The argument behind not training a deep convolutional network is a strawman argument.
There is a also learning stage inherent to our approach: LDA or LMNN. @writing}
\ml{so argue the lack of large dataset ? Or not talk about it}

\vl{I do believe that time-frequency scattering (or one of its variants, see McDermott or Patil or Sch\"{a}dler)
is, in a very precise way, mandatory to model the central auditory system. It's OK to think otherwise but that
should be motivated. @writing}
\ml{then you can proceed to the rewrite}

\vl{I would suggest trying to avoid the invention of the "TST" acronym for "temporal scattering transform" if possible.
This acronym does not already exist in the relevant literature. @writing}
\ml{this is not an invention, this is a mean to be compact in this paper.}

@@

\vl{I am a bit wary about not giving proper names to the two experiments, but instead calling them "experiment 1" and "experiment 2".
How about "expert judgments of uniqueness" versus "crowdsourced judgments of similarity"? @writing}
\ml{J'ai bien peur que cela alourdisse le discours. Mais bon, visiblement on ne repete experiment 2 que 6 fois, ca devrait aller.}

@@

\vl{The paragraph above is the most important in the whole paper, and it deserves more attention. How do these three solutions differ? Which one was chosen in the end? Did it make a significant difference in NMI? Is the consensus clustering shown somewhere in the paper? Is it related to the dendrogram in Appendix B? I have so many questions. @writing}
\ml{ok, je refais un code sur cette partie. Le dendrogram n'est pas un clustering, il part de la matrice de similarité. Bon apres, c'est tres lié...}

\vl{I'm afraid there is also a methodological problem here.
We need to have a hold-out set of IPTs to evaluate the performance of the clustering algorithm after having trained the metric learning algorithm.
@methodology @computation}
\ml{Mmm, je ne suis pas sur, dans l'idée, j'ai besoin d'une projection adaptée a mon contenu, rien ne m'empeche d'optimiser sur tout le corpus. On peut imaginer couper en deux mais ca fait pas beaucoup de données. Perso, je suis pour laiser ca comme ca et le faire si un relecteur rale. Mais je suis ouvert a la discussion.}

@@

\vl{What does "safely" mean here? We are at the epicenter of the paper here: every word matters. @writing}
\ml{I agree, we should develop a bit.}

@@

As explained in the first section, gathering perceptual similarity data is plagued with a size problem. Thus, the perceptual data considers 78 \ipts each represented by one sound recording, most often of C4 pitch and nuance mezzo forte. In order to provide enough data for the numerical analysis described in this section, we complement the data gathered during the second experiment as follows: for each \ipt, instead of one sound exemplar we consider the whole range of nuance and pitch that is available. This leads to a dataset of 9346 sound recordings. It is thus assumed that the perceptual similarity is invariant with respect to pitch and nuance which is not necessarily true, but is taken in this paper as a reasonable assumption.
\vl{It's too bad that this "reasonable" assumption is not challenged against the data. In music psychology, the "bandwidth for timbre invariance" is of the order of one octave (Handel and Erickson, Music Perception, 2001) or less (Steele and Williams, Music Perception, 2006).
One way to do it without collecting new similarity judgments would be to break down the precision by pitch of query.
Presumably we have an inverted U shape with a peak near C4.
@research @computation @figure @biblio}
\ml{Indeed, It would be interesting but would require some research time.}

@@

\vl{I'm extremely puzzled by the paragraph above. I thought that the point of the paper was precisely to liberate oneself from the taxonomy of musical instruments? @writing}
\ml{This is just an example. May be state better that is htis example the reference is the instrument taxonomy. Or even better refer to a clustering reference of one subject, but in this case the cluster denomination is more cryptic, like cluster 1, cluster 2...}

@@

\vl{The mel scale has a varying quality factor, so I don't see how this can be right.}
\ml{I don't understand}
\vl{I understand what "logarithmically compressed" and "normalized using the median" because I did work on this before but,
for someone who is less experienced, this might not be precise enough.}
\ml{ok}

@@

\vl{The paragraph above is very peremptory and at the same time quite vague.
I could quote many tasks in which there is no need to "target in some way" the scattering transform to obtain better results than MFCC.
What scientific conclusion are we drawing here? @writing}
\ml{to be discussed}

@@

\begin{table}
  \caption{Precision achieved by the different settings. Numbers in bold face indicate best performance. \vl{What does these variances correspond to? Variance across raters, across IPTs, across instances of each IPT? Furthermore, why are the results with TST, LMNN, and "reference" commented? Do they correspond to same rater between training and evaluation? @methodology @writing}
  \ml{variance across raters. Do not understand the second question, to be discussed.}}
  \label{tab:res1}
  \begin{center}
\begin{tabular}{lllc}
features & projection & reference & precision (\%) \\
  \hline
  MFCCs & - & - &   85.07 $\pm$ 6.19 \\
  MFCCs & LDA &  consensus  &   81.50 $\pm$ 7.65 \\
MFCCs & LMNN & multiple & 86.31 $\pm$ 5.91 \\
MFCCs & LMNN & consensus &   86.22 $\pm$ 5.92 \\
% MFCCs & LMNN & reference &   86.18 $\pm$ 6.05 \\
TST & - & - &   87.01 $\pm$ 5.81 \\
TST & LDA & consensus &   80.95 $\pm$ 10.37 \\
TST & LMNN  & multiple  &   93.31 $\pm$ 3.92 \\
TST & LMNN & consensus &   \textbf{94.80 $\pm$ 3.26} \\
% TST & LMNN & reference &   98.09 $\pm$ 1.28 \\
\end{tabular}
\end{center}
\end{table}

@@

\begin{figure}
\center
\includegraphics[width = \textwidth]{figures/frame.png}
\caption{Evolution of the precision with respect to the frame size for the different front ends.
\vl{The x axis should be made logarithmic, which uneven tick marks.
The legend is difficult to understand (acronyms, subtle difference between "multiple" and "consensus").
Again, the legend box could be removed after placing the labels next to the lines themselves. @figure}}
\ml{Proposition ?}
\label{fig:frame}
\end{figure}

@@

For the sake of reproducibility, the code of the proposed method and the experimental protocol is available\footnote{https://github.com/mathieulagrange/paperSpontaneousSimilarity \vl{I would love a more memorable name. @writing}\ml{proposition ? abbreviation of the new title ?}}. For convenience, the features (scattering and MFCCs) are also provided, but the software can be used to process audio data.

@@

Thanks to the digitization capabilities available, gathering large amount of audio recordings of \ipt is achievable and several databases are now available on the market. Browsing of such databases can be achieved by keyword-based search. It requires that 1) the database is fully annotated with a precise ontology of instruments and playing techniques, 2) that the users have a precise understanding of this ontology, and 3) that the users a precise idea of the kind of timbre they are looking to find the \ipt they are searching for.

\vl{Curiously, this paper does not contain a review of such databases available on the market. @biblio}
\ml{Indeed}

@@

Some instruments can be complemented with sordina (S), wha (W). The addition of this kind of device modifying the shape of the instrument, a saxophone alto and saxophone alto with sordina are considered as different instruments.
\vl{This is crucial and should not be deferred to the appendix. @writing}
\ml{J'ai mis ca en appendice pour ne pas que le lecteur confonde la base SOL complete et les données considérées dans l'étude.}

@@

\vl{This section is not precise enough in terms of intensity ranges and pitch ranges.
The abbreviations can be removed. @data @writing}
\ml{}

\vl{A bubble plot of number of notes per instrument (x axis) and playing technique
(y axis) would be extremely useful. @data @figure}
\ml{Tu voudrais ca pour la base SOL complete ou les données considérées dans l'étude ?}

@@

\vl{This dendrogram is very hard to read. Abbreviations must be replaced by the full names of every instrument and playing techniques.
Another way to visualize it would be a two-dimensional embedding, such as t-SNE or Isomap (I'm biased towards Isomap). @figure}
\ml{Isomap do not show a dendrogram, right ?}
